\chapter*{Curriculum Vit\ae}
\addcontentsline{toc}{chapter}{Curriculum Vit\ae}
\setheader{Curriculum Vit\ae}

%% Print the full name of the author.
\makeatletter
\authors{\@firstname\ {\titleshape\@lastname}}
\makeatother

\noindent
\begin{longtable}{p{.225\textwidth} p{.70\textwidth}}
    1993/04/07 & Date of birth in Tehran, Iran
\end{longtable}

\section*{Education}
\begin{longtable}{p{.225\textwidth} p{.70\textwidth}}
    10/2019-06/2024 & Ph.D. Student, Software Engineering Research Group (SERG), \\
     & Delft University of Technology, The Netherlands, \\
     & Thesis: Machine Learning-assisted Software Analysis, A \\
     & Supervisors: Dr. Sebastian Proksch (2021-2024) \\
     &  \hspace*{5.1em} Dr. Georgios Gousios (2019-2020) \\
     & Promotor: Prof. Dr. Arie van Deursen \\
     & \\

     01/2016-01/2019 & M.Sc., Computer Engineering (Artificial Intelligence), \\
     & Azad University (North Tehran Branch), Tehran, Iran, \\
     & Thesis: Robust Twin Support Vector Machine for Noisy Data, 4/4 \\
     & Supervisor: Dr. Jalal A. Nasiri \\
     & \\

     10/2011-07/2015 & B.Sc., Computer, University of Tehran, Tehran, Iran \\
     & Thesis: Applications of the Python Programming Language for Climatology, 3.95/4\\
     & \\
     07/2010-07/2011 & Pre-university Diploma, Alavi, Tehran, Iran \\
     10/2007-06/2010 & High School Diploma, Andisheh, Tehran, Iran \\    
\end{longtable}

\section*{Experience}
\begin{longtable}{p{.225\textwidth} p{.70\textwidth}}
    10/2019-09/2022 & Research Software Engineer, the FASTEN Project, \\
    & Delft University of Technology, The Netherlands \\
    & \\
    07/2017-09/2019 &  Research Assistant, Machine Learning and Text Mining Lab \\
    & Iranian Research Institute for Information Science (IranDoc), Tehran, Iran \\
\end{longtable}

\newpage
\section*{Academic Service}
\begin{longtable}{p{.225\textwidth} p{.70\textwidth}}
    Chair & Social Media and Publicity Co-chair MSR 2024 \\
    & \\
    PC Member &  ASE, Industry Track, 2024 \\
    & ICSE 2024, Artifact Evaluation \\
    & NLBSE 2024 \\
    & MSR 2023 (Junior PC) \\
    & \\

    Reviewer & ICLR 2023 \\
             & JSS 2023 \\
             & TOSEM 2023 \\
             & Expert Systems with Applications, 2021\\
             & FSE 2021, Artifact Evaluation (external) \\
             & \\
    
    Council Member & PhD council of the Institute for Programming \\
                  &  Research and Algorithmics (IPA), 2021-2024 \\
                  & \\

    Student Volunteer & ICSE 2024 \\
    & \\

    Virtualization (Live Stream) & ICSE 2020 \\
    & \\
    Co-supervision & Lang Feng's Master Thesis "Static Analysis Complements Machine Learning: A Type Inference Use Case", 2022-2023 \\
    & Jonathan Katzy's Master Thesis "Utilizing Lingual Structures to Enhance Transformer Performance in Source Code Completions", 2021-2022 \\
    & Bachelor end project of a group of three students studying the combination of static analysis with deep learning to improve type inference for Python, 2022 \\
    & \\
    Teaching Assistant & Release Engineering for Machine Learning, Dr. S. Proksch \& Dr. L. Cruz, 2022 \\
    & Machine Learning for Software Engineering, Dr. Georgios Gousios, 2020 \\ 
\end{longtable}

