\chapter*{List of Publications}
\addcontentsline{toc}{chapter}{List of Publications}
\setheader{List of Publications}
\label{publications}

%% We use the 'etaremune' environment (the reverse of 'enumerate') to get a
%% numbered list of publications in reverse chronological order. If the list of
%% authors is long, it might be useful to emphasize your own name with \textbf.
\begin{etaremune}{\small
\item[\faFileTextO~~1.] \textbf{Amir M. Mir}, Mehdi Keshani, and Sebastian Proksch. On the Effectiveness of Machine Learning-based Call Graph Pruning: An Empirical Study. In IEEE/ACM 21st International Conference on Mining
Software Repositories (MSR'24), April 2024

\item[\faFileTextO~~2.] \textbf{Amir M. Mir}, Mehdi Keshani, and Sebastian Proksch. On the Effect of Transitivity and Granularity on Vulnerability Propagation in the Maven Ecosystem. In Proceedings of the 30th IEEE International
Conference on Software Analysis, Evolution and Reengineering (SANER'23), March 2023

\item[\faFileTextO~~3.] \textbf{Amir M. Mir}, Evaldas Latoskinas, Sebastian Proksch, and Georgios Gousios. Type4Py: Practical Deep Similarity Learning-based Type Inference for Python. In Proceedings of the 44th International Conference on
Software Engineering (ICSE'22), pages 2241–2252, May 2022

\item[\faFileTextO~~4.] \textbf{Amir M. Mir}, Evaldas Latoskinas, and Georgios Gousios. ManyTypes4Py: A Benchmark Python Dataset for
Machine Learning-Based Type Inference. In IEEE/ACM 18th International Conference on Mining Software Repositories (MSR'21), pages 585–589, May 2021

\item[5.] Venkatesh, A. P. S., Sabu, S., Wang, J., \textbf{Mir, A. M.}, Li, L., \& Bodden, E. TypeEvalPy: A Micro-benchmarking Framework for Python Type Inference Tools. Tool Demonstration, In Proceedings of 46th International Conference on Software Engineering (ICSE'24), April 2024

\item[6.] Venkatesh, A. P. S., Sabu, S., \textbf{Mir, A. M.}, Reis, S., \& Bodden, E. (2024). The Emergence of Large Language Models in Static Analysis: A First Look through Micro-Benchmarks. First special event of AI Foundation Models and Software Engineering (FORGE'24), April 2024
}
\end{etaremune}

\vspace{0.5cm}
\noindent
\faFileTextO~~Included in this thesis.\\
% \faTrophy~~Won a best paper, tool demonstration, or proposal award.
